% This is "sig-alternate.tex" V1.9 April 2009
% This file should be compiled with V2.4 of "sig-alternate.cls" April 2009
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.4 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.4) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
%  ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V1.9 - April 2009

\let\mymarginpar\marginpar

\documentclass{sig-alternate}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{array}
\usepackage{booktabs}
\usepackage{capt-of}
\usepackage{colortbl}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{pifont}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{url}
\usepackage{boxedminipage}
\usepackage{xspace}
\usepackage{sepnum}
\usepackage{cite}
\usepackage{fancyhdr}

% Things to remember
\newenvironment{mynote}%
{\medskip
\noindent
\vspace{2pt}
%\setlength{\fboxsep}{4pt}
%\setlength{\leftmargini}{4pt}
\let\emph=\textbf
\begin{boxedminipage}{.99\columnwidth}\em}%
{\end{boxedminipage}}

\newcommand\remark[1]{%
    \mymarginpar{\raggedright\hbadness=10000\tiny\it #1\par}}
\def\todo#1{\noindent\colorbox{yellow}{X}\remark{{\bf TODO:} #1}}
\def\fixme#1{\noindent\colorbox{yellow}{X}\remark{{\bf FIXME:} #1}}
\def\TODO#1{\todo{#1}}
\def\FIXME#1{\fixme{#1}}

\newcommand\tm{\texttrademark}

\overfullrule3pt

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

%\pagestyle{fancy}
%\fancyhead{Submitted to ICSE 2009}


%\title{STC and Build Success in Jazz}
%\title{Is NOT Talking About Your Impact on Others Harmful?}
\title{On Socio-Technical Coordination and its Relation \\to Build Failure}

%\title{Communication Patterns}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{2} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Adrian Schr{\"o}ter\\
       \affaddr{Univeristy of Victoria, Canada}\\
       %\affaddr{1932 Wallamaloo Lane}\\
       %\affaddr{Wallamaloo, New Zealand}\\
       \email{schadr@uvic.ca}
% 2nd. author
\alignauthor
Daniela Damian\\
       \affaddr{University of Victoria, Canada}\\
%       \affaddr{P.O. Box 1212}\\
%       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{danielad@cs.uvic.ca}
%% 3rd. author
%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
%one who did all the really hard work.}\\
%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
%       \affaddr{Hekla, Iceland}\\
%       \email{larst@affiliation.org}
%\and  % use '\and' if you need 'another row' of author names
%% 4th. author
%\alignauthor Lawrence P. Leipuner\\
%       \affaddr{Brookhaven Laboratories}\\
%       \affaddr{Brookhaven National Lab}\\
%       \affaddr{P.O. Box 5000}\\
%       \email{lleipuner@researchlabs.org}
%% 5th. author
%\alignauthor Sean Fogarty\\
%       \affaddr{NASA Ames Research Center}\\
%       \affaddr{Moffett Field}\\
%       \affaddr{California 94035}\\
%       \email{fogartys@amesres.org}
%% 6th. author
%\alignauthor Charles Palmer\\
%       \affaddr{Palmer Research Laboratories}\\
%       \affaddr{8600 Datapoint Drive}\\
%       \affaddr{San Antonio, Texas 78229}\\
%       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
%Investigating collaborative aspects of software development is becoming 
%prominent in current research. Studies found that the misalignment
%between the social and technical dimensions of software work leads to losses in
%developer productivity and defects. We study the communication and technical dependencies between developers involved in software builds and
%relate their misalignment to the integration success. Using data from the
%IBM Jazz\texttrademark\ project we investigate socio-technical networks in relation to build failure. 
%Our findings related a number of
%developer pairs that did not communicate about their dependencies to build failure.
%Not only can we use these pairs to predict build failure but, moreover, it is actionable knowledge upon developers and mangers can act to prevent build failure.  
%If any one of these pairs is present in a
%social network of a build, the build had at least an 74\% chance to fail.
%
Investigating social interactions in software development is becoming 
prominent in current research. The misalignment
between the social and technical dimensions of software work has been
linked to losses in developer productivity and defects. In a case study of
coordination in the IBM Jazz\texttrademark\ project, we investigate the
communication and technical dependencies between developers involved in
software builds and relate their misalignment to the build failure. We found
that historical project information about socio-technical coordination and software
builds can be used in a model that predicts the quality of upcoming builds. We
also identify a number of
developer pairs that did not communicate about their dependencies and thus
increased the likelihood of build failure. Upon this actionable knowledge developers and mangers can act to prevent build failure. If any one of these pairs is present in a
builds social network, the build had at least an 74\% chance to fail. 
This has several practical implications for the design of collaborative systems, such as the integration of recommendations about inter-personal relationships.
%
%We discuss the practical implications for the design of collaborative systems that
%can integrate real-time recommendations about the inter-personal relationships
%that matter to avoid build failures in the project. 
\end{abstract}


\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, \linebreak performance measures]
\category{D.2.9}{Software Engineering}{Management}[Programming Teams]
\category{K.6.1}{Management of Computing and Information Systems}{Project and People Management}[Systems development]
\category{K.6.3}{Management of Computing and Information Systems}{Software Management}[Software development]

\terms{Human Factors, Measurement, Management}

\keywords{Social Networks, Technical Networks, Socio-Technical Networks, Builds, Failures, Socio-Technical Congruence}

%\pagebreak
\section{Introduction}
We hypothesize that with the ever growing size of software teams the lack of
effective coordination is the main source of integration failures.  With the
ever growing complexity and sophistication of large software projects,
error-free integrations are not only important but difficult to achieve. The
development work that precedes integrations involves significant coordination of
developers that work in teams and need to rely on the code of others and its
stability. But often code is everything but stable, further contributing to
developers' need to coordinate to keep up with code changes that impact their work. This problem is
amplified in software builds where an entire team needs to integrate their work
and on which the development of new features depends. Not only do
failed builds destabilize the product~\cite{cusumano1997} but they also demotivate
software developers~\cite{holck2004}.

Despite their importance, keeping integrations builds error-free can be a very time consuming
process. A lightweight approach that can determine whether the build contains
failures before invoking the build process is thus very valuable to developers.
This lightweight approach could determine a builds outcome in minutes rather than hours or days. Having a faster way to
assess the quality of a build helps developers to continue working with newest
builds while being aware of its quality. Previous
research~\cite{wolf:icse:2009,hassan:ase:2006} trained predictive models to assess the quality of software builds without the need of invoking large test
suits. Although this research
reaches a high degree of accuracy in their predictions, knowing that a
build will fail does not necessarily help developers to actually prevent
the build from failing.
% limitation of previous research
% SENTENCE HERE ABOUT WHY THIS IS A PROBLEM. SENTENCE ABOUT HOW THIS WAS STUDIED
% BEFORE AND THE LIMITATION THAT RESEARCH HAD ONLY STUDIED PRECTIVE MODELS
% WITHOUT THE ABILITY TO ACT ON THAT KNOWLEDGE.
The goal of this research is to find a way to create actionable knowledge that
developers can act upon to avoid
integration failure.

In this paper we describe a case study of IBM's Jazz project where we leverage
information about socio-technical developer coordination and software builds to
identify pairs of developers that negatively influence the quality of the
upcoming build. Using historical project information we construct socio-technical
networks that capture information about developers with technical dependencies
as well as their ongoing communication as conceptualization of their coordination.
We found that using a support vector machine on a combination of this social
and technical project data yields a powerful predictor of build failure. We
then identify that there are certain pairs of developers that have a
negative influence on the build outcome. On this knowledge developers and management can
act upon to avoid future build failure.

Maintaining proper communication and awareness of work others perform is
important in any kind of project. Specifically in software engineering many studies found
that factors such as geographical and organizational distance have an impact on
communication and even effect software quality~\cite{nagappan:icse:2008}. In our
study we uncover the existence of pairs of
developers, that, if technically dependent in a build but not discussing their
dependencies, have a negative influence on the success on builds. This
actionable knowledge can be integrated in real-time recommender systems that
indicate, based on project historical data, which developer pairs tend to be
failure related. Developers and management can then devise strategies to
prevent the failure before build time. 


The paper is organized as follows: We start with formulating our research
questions (Section~\ref{sec:rq}) followed by putting our research into the
perspective of related work (Section~\ref{sec:relwork}). Section~\ref{sec:data}
covers details about our case study of IBM's Jazz team to investigate
socio-technical coordination in relation to builds. Then
Sections~\ref{sec:prediction} and \ref{sec:pattern} cover the actual analysis and
their discussion. Before we concluding the paper in
Section~\ref{sec:conclusions}, we discuss some practical implications in Section~\ref{sec:implications} and the threats to validity
(Section~\ref{sec:threats}).



%  ONCE SENTENCE HERE REMINDING THE READER THAT IN THIS PAPER WE PROPOSE A
% LIGHTWEIGHT APPROACH THAT LEVERAGES DEVELOPERS SOCIAL NETWORKS BUILT FROM
% PROJECT HISTORICAL DATA AND WHICH IDENTIFIES WHICH DEVELOPER PAIRS ARE
% IMPORTANT WITH RESPECT TO THE QUALITY OF THE UPCOMING BUILD.

\section{Research Questions}
\label{sec:rq}
Communication
is an important mechanism for coordination in software development \cite{curtis:acm:1988,kraut1995:coordination}, especially in maintaining the awareness of changes relevant to your work. Past research showed that a high coverage of technical
dependencies by social relationships correlates with higher productivity of development teams (e.g.~\cite{cataldo:cscw:2006}). 
 We hypothesize that a similar relation
exists between this coverage and the success of software builds. 

Thus
socio-technical networks that capture information about the social and technical
relationships between developers, should predict build failure:

\begin{itemize}
\item[RQ1] Can we use information about relations (technical and social) between
developers to predict build outcome?
\end{itemize}

Although valuable, having a model that predicts build outcome is often not
enough. We seek to generate actionable knowledge upon which developer can act to
avoid a build from failing. Past research suggests that the absence of
communication between developers that are technically dependent leads to
problems, such as slow down in development~\cite{cataldo:esem:2008}.


We hypothesize that due to the high coordination needs the absence of this
important communication also has a negative influence on build outcome. Communication problems can arise from many factors including organizational,
social or technical reasons. Being able to pinpoint
mismatches between technical dependencies and required communication that
relate to build failure is even more important in a team's
ability to devise strategies to avoid build failure. Thus, we
investigate pairs of developers that share a technical dependency without talking
with each other (referred to as \emph{technical pairs}):


\begin{itemize}
\item[RQ2] Are there technical pairs that influence the build outcome?
\end{itemize}

Acknowledging that build failure can be the result
of factors other than lack of developer
communication, e.g. the number of changes or developers in a build~\cite{hassan:ase:2006}, our
analysis also studies the effect of technical pairs in the presence of such
confounding variables. 



%The answer to this question will tell us if we really can devise strategies to
%prevent builds from failing. 


\input{relatedwork}
%\pagebreak
\input{methodology}

\input{failure-prediction}

\begin{table}[t]
\centering%\vspace{1cm}
\begin{tabular}{rcc}
\toprule
 & successful & failed  \\
 \midrule
(Adam, Bart) & 3 & 13 \\
$\neg$ (Adam, Bart) & 224 & 86\\
\midrule
total&227&99\\\bottomrule
%User3493, User2943 & 3 & 13
\end{tabular}
\caption{Contingency table for technical pair (Adam, Bart) in relation to build
success or failure}
\label{tab:contingencytable}
\end{table}

\input{pattern-analysis}

%\input{regression-analysis}







\section{Practical Implications}
\label{sec:implications}
Our findings have several implications for the design of collaborative systems.
By automating the analyses presented here we can incorporate the knowledge about
developer pairs that tend to be failure related in a real-time recommender
system. Not only do we provide the recommendations that matter to the upcoming
build, we also provide incentives to motivate developers to talk about their
technical dependencies. 

Such a recommender system can use project historical data to
calculate the likelihood that an upcoming build fails given a particular
developer pair that worked on that build without communicating to each other. In
the case of the pair (Adam, Bart) the system may recommend that these
developers should communicate about their technical dependencies, as there
would be a probability
of 91\% of failure of the next build should they not follow the
recommendation. This probability of failure serves as mechanism to
rank importance of a socio-technical gap and more importantly as an
incentive to act upon.

For management, such a recommender system can provide details about the
individual developers in, and properties of, these potentially problematic
developer pairs. Individual developers may be an explanation for the behavior of
the pairs we found in Jazz\texttrademark. This may indicate developers that are
harder to work with or too busy to coordinate appropriately, prompting management
to reorganize teams and workloads. This would minimize the likelihood of a build
to fail, by removing the underlying cause of a pair to be failure related.
Similarly, as another example from our study, most developer pairs
consisted of developers that were part of different teams. In such
situations management may decide to investigate reasons for coordination
problems that include factors such as geographical or functional distance in the project.




\section{Threats To Validity}
\label{sec:threats}
%\todo{rework}
During our study we identified two main threats. 
One threat covers issues that arise from the underlying data we used.
The other threat deals with possible problems from the conceptualization of
constructs in our study.

\subsection{Data}
We performed all our analysis on one set of data, the Jazz\texttrademark\
repository. This limits the generalizability of our findings, due to the fact
that we only made the observations within one project. The  
project size and the project properties -- incorporating open
source practices such as open development and encouraging community involvement
-- make us believe that our findings still hold value.

Furthermore we only investigated three months of the project's lifetime. This
might lead to smaller significance of our results. However, since the three
months are directly before a major release of the project, this dataset contains
the most viable data for our analysis. In those three months a lack of necessary
coordination is the most harmful to the project.

Another threat that is inevitable in studies of software engineers is
the possible lack of recorded communication. This and the possibility of people
coordinating without communicating, such as reading each others source
code~\cite{bolici:stc:2009}, are mitigated in Jazz by its development process.
The Jazz\texttrademark\ team's development process demands that the developers
coordinate using workitem discussion. 


\subsection{Conceptualization}
Our conceptualization of the three edges we use to construct the
socio-technical networks might introduce inaccuracy in our findings.
First, social edges are extracted from workitem discussions. We assumed that
every developer commenting on or subscribed to a work item reads all comments of that
work item. This assumption might not always be correct. By manual inspection of
a selected number of work items, however, we found that developers who
commented on a work item are aware of the other comments, confirming our assumption.

Second, the technical edges are not problematic by themselves, but they are not
complete, since there are more technical relationships between developers that
can be examined. For example, two developers can be connected if one developer
changes someone else's code. This however does not invalidate our findings, it
just suggests that there is room for improvement, which we should address
next.

Third, socio-technical edges on the other hand may suffer from the combination of
social and technical edges. For example, it is not necessarily true that the
discussion of  two developers in a technical dependency is always about their
technical dependency. 
In our study however, since the changes to
source code files we use to extract technical dependencies are attached to workitem discussion, we are
confident that they addressed the changes at least indirectly.



\section{Conclusion}
\label{sec:conclusions}
Our study investigated the relationship between pairs of developers that share
a technical dependency but do not communicate and build failures. We were
motivated by findings in the literature that suggested that high alignment between technical dependencies and actual coordination in a project has a positive effect on task
performance.
%
We hypothesized that a similar relationship may be found in relation to a broader
coordination outcome, i.e. integration outcome, because developers not
coordinating about dependencies in their work might lead to errors remaining in
the code that break the build.

Our case study of coordination in the Jazz project indicates that 
historical project information about socio-technical coordination and software
builds can be used in a model that predicts the quality of upcoming builds. We also found
that the influence of developer pairs with a socio-technical gap on the build
failure was very high. This means that if any one of these pairs was present in a
social network of a build, the build had at least an 74\% chance to fail.
Thus, we add to the literature on failure prediction with a method that
provides actionable knowledge to avoid failures. 

Besides the possible of integrating our findings into a real-time recommender system that indicates the specific developer pairs that should talk about their dependencies, our results add to the research in
collaborative software engineering. Not only do we add to the evidence about
the role of effective communication in the development of high quality
software, but our findings indicate that measures of socio-technical congruence can
serve as a mechanism to identify which inter-personal relationships are
important to avoid breaking software integrations. 
%
We plan to extend this research in a number of directions:

% \begin{description}
\emph{More specific information in socio-technical networks.}
We plan to study socio-technical coordination at a finer level by refining all
three edge types social, technical and socio-technical as well as the
information about developers. For the social edges we plan on identifying who
people are talking to and exactly about what. Technical edges can be refined by examining other source code relations, such as call
graphs, or changes made to others' source code. 
%
To combine social and technical
edges to socio-technical edges we plan to use content analysis techniques on
communication to match it to the appropriate technical edge. We also plan
on also investigating the developers that are part of the social networks more
closely by incorporating developer characteristics, such as experience,
geographical location, role or team allocation.

\emph{Additional edges that indicate emergent coordination.}
In this study we focused on technical pairs and only briefly touched on
socio-technical pairs. In the future we plan to extend this focus to include a
detailed analysis of socio-technical pairs and of those
developer pairs that communicated without sharing a technical dependency, and
which indicate emerging communication in the project, and expertize seeking
behaviors that are important to effective coordination.

\emph{Different types of failures.}
A build fails often because of a single or several failures scattered across several locations
Hence we plan on
investigating single bugs. This additionally enables us to investigate pre-release and post-relase failures separately.
%\end{description}


%\pagebreak
\section{Acknowledgements}
This project is funded by an IBM Jazz Innovation Award and a Univeristy fellowship of the University of Victoria.
Thanks for continuous feedback go to the SEGAL Group members especially to Irwin
Kwan, Sarbina Marczak, and Mattew Richards.
And special thanks goes to the Jazz\tm\ development team as well as the IBM
research for they ongoing support.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{fse2010-schroeter}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns

%\balancecolumns % GM June 2007
% That's all folks!
\end{document}
