\documentclass[conference]{IEEEtran}

\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
   \graphicspath{{../pdf/}{../jpeg/}}
\else
\fi

\usepackage[cmex10]{amsmath}

\usepackage[tight,footnotesize]{subfigure}

\usepackage{booktabs}

\begin{document}

%\title{Deriving Actionable Knowledge from Information on Socio-Technical Gaps in Software Projects}

\title{Deriving Actionable Knowledge from Socio-Technical Gaps in Developer Coordination}

%On Socio-Technical Coordination and its Relation to Build Failure}


\author{\IEEEauthorblockN{Adrian Schr{\"o}ter}
\IEEEauthorblockA{University of Victoria, Canada\\
schadr@acm.org}
\and
\IEEEauthorblockN{Daniela Damian}
\IEEEauthorblockA{University of Victoria, Canada\\
danielad@cs.uvic.ca}}

\maketitle

% MKAE IT TALK ABOUT THE APORACH AND NOT CASE STUDY

\input{abs}

\IEEEpeerreviewmaketitle

\section{Introduction}
A growing body of research (e.g. ~\cite{kwan:tse:2011,herbsleb:icse:1999}) shows that with the ever growing size of software teams the lack of
effective coordination is the main source of integration failures. The
development work that precedes integrations involves significant coordination of
developers that work in teams and need to rely on the code of others. 
Unfortunately, code is often anything but stable, further contributing to developers' need to coordinate with one another and stay up to date with code changes that impact their work.
%But often code is everything but stable, further contributing to
%developers' needing to coordinate to keep up with code changes that impact their work. 
This problem is
amplified in software builds that integrate an entire team's work
and on which the delivery of new features depends. Not only do
failed builds destabilize the product~\cite{cusumano1997}, they also demotivate
software developers~\cite{holck2004}.

%Keeping integrations builds error-free can be a very time consuming
%process. 
One major challenge with integrating work across teams is coordination~\cite{cataldo:esem:2008}.
Coordination problems can arise from factors of organizational,
social or technical nature~\cite{herbsleb:icse:1999}. Studies showed that the mismatch between technical dependencies and required communication to tackle these dependencies increases the likelihood of integration failures~\cite{kwan:tse:2011}. Similarly, research~\cite{wolf:icse:2009,hassan:ase:2006} trained predictive models to assess the quality of software builds without the need to invoke large test suits. Although this research
reaches a high degree of accuracy in their predictions, knowing that a
build will fail or that failures happen due to socio-technical mismatches does not necessarily help developers to prevent the build from failing. Pinpointing more precise build-related coordination problems enhances a team's ability to devise strategies to avoid build failure. 



%relate to build failure is even more important for a team's
%ability to devise strategies to avoid this problem.
%Thus an approach to support developers to identify co-workers they need to coordinate with in order to avoid the up coming build from breaking is called for.
%Thus it is important to initiate an approach that supports developers in identifying coworkers that they need to coordinate with in order to avoid an up coming build from breaking.


%The goal of this research is to find a way to create actionable knowledge to avoid
%integration failure.

Our research seeks to identify actionable knowledge to avoid integration failure by providing developers with information with which coworkers they need to coordinate with in order to avoid failure of the upcoming build. In this paper we describe an approach that we evaluated with data from the IBM's Rational Team Concert project in which we leverage
information about socio-technical developer coordination and software builds to
identify pairs of developers that increase the likelihood of a build failing. From historical project information we extract information about developers with technical dependencies
as well as their ongoing coordination. From this, certain pairs of developers that have a negative influence on the build outcome are identified. This
actionable knowledge can be integrated in real-time recommender systems and developers and management can
devise strategies to prevent failure before build time.

%Through the application of our proposed approach we uncover the existence of pairs of
%developers that if technically dependent in a build but not discussing their
%dependencies, have a negative influence on build success. 

Our approach is based on the hypothesis that due to the high coordination needs in a build the absence of corresponding coordination increases the likelihood of build failure. 
Thus, we
investigate pairs of developers that share a technical dependency in the absence of actual coordination (referred to as \emph{technical pairs}) and formulate our research question:

\textbf{RQ} How can we identify developer pairs that negatively influence build success?
 

\section{Related Work}
\label{sec:relwork}
We aim to integrate work that investigates team collaboration to produce actionable knowledge upon which developers can act.
Several studies are relevant with respect to different dimensions of our work:

\paragraph{Factors that affect software builds}
To the best of our knowledge the studies by Hassan et al.~\cite{hassan:ase:2006}
and Wolf et al.~\cite{wolf:icse:2009} are the only studies that conducted
research to predict build outcome. Hassan et al.~\cite{hassan:ase:2006} found
that a combination of social metrics (e.g. number of authors) and technical
metrics (e.g. number of code changes) derived from the source code repository
yield to be best predictor. 
On the other hand Wolf et al.~\cite{wolf:icse:2009} showed that communication structure has an influence on the build outcome.
Kwan et al~\cite{kwan:tse:2011} investigated socio-technical networks with an emphasis on technical dependencies among developers that is not accompanied with any coordination activity on build outcome (socio-technical gap).
We complement the work of Kwan et al's ~\cite{kwan:tse:2011} that showed a relationship between the number of gaps and build failure, by leveraging information about these gaps to create actionable knowledge.

\paragraph{Coordination in software development}
In order to manage changes and maintain quality, developers must coordinate. In
software development, coordination is largely achieved by communication with those sharing work dependencies~\cite{kraut1995:coordination}. 
The software engineering literature is recognizing the role of communication as essential~\cite{nakakoji2010:rdc}.
Several researchers in the software engineering community investigated the effects of communication on several topics such as knowledge distribution~\cite{ehrlich:icgse:2006}, coordination~\cite{hinds:cscw:2006}, and Conway's Law~\cite{cataldo:cscw:2006}.
We build onto these findings in devising an approach that improves team communication.


\paragraph{Coordination and failure in software development}
Many studies bring evidence of the relationship between coordination and software integrations. Besides qualitative studies (e.g~\cite{herbsleb:icse:1999}), quantitative approaches often represent coordination among developers in the form of social networks. These social networks, when related to actual code artifacts, can be used to predict the artifacts' failure-likelihood.
Several studies showed that metrics derived from social networks form good failure predictors (e.g.~\cite{meneely:fse:2008}).
Furthermore, when combining these social networks with information of organizational hierarchies~\cite{nagappan:icse:2008}, or geographical distance~\cite{bird:acm:2009}, they yield not only better predictors but also shine a light on other factors that influence developers' coordination.
% herbsleb grinter
%Herbselb and Grinter~\cite{herbsleb:icse:1999} found that coordination impacts integration efforts.
%We take the findings that show the influence of coordination on software quality and combine them with technical dependencies among developers to generate actionable knowledge.

Recent work has studied the effects of socio-technical alignment on coordination and project outcomes. The mismatch between coordination needs and actual coordination -- implied by work dependencies and developer social interactions respectively -- is referred to as a socio-technical gap. Empirical studies related socio-technical gaps to low developer productivity~\cite{valetto:msr:2007} and low software quality~\cite{kwan:tse:2011}.
With respect to software builds, Kwan et al~\cite{kwan:tse:2011} investigated the occurrence of socio-technical gaps throughout builds in the Rational Team Concert project highlighting that builds that contain more gaps are more likely to fail.
%Valetto et al~\cite{valetto:msr:2007} uncovered a similar insight into productivity related to implementing work items showing that the existence of gaps is a major cause for low productivity.
These findings suggest that the mismatch between implied coordination needs and actual coordination should be avoided to both guarantee high productivity and software quality.
  

%MOVE THOSE DOWN
%A gap is represented by two developers that share a technical dependency (implying coordination need) without any social interaction (implying
%unmet coordination need).
%In this paper, we refer to these pairs of developers as \emph{technical pairs} (there is a gap), and to those that do
%share a socio-technical dependency (there is no gap) as \emph{socio-technical pairs}.


%MOTIVATE THE APPROACH
%
%WHAT IS THE GOAL OF THE APPROACH
%
%WHAT TYPE OF PROJECTS ARE YOU APPLYING THIS TO
%
%EXAMPLE FOR TECHNICAL PAIRS, HOW DO I BUILD THOSE

\section{Formulating the Approach}
Motivated by research showing that socio-technical gaps increases the chance of a build to fail we formulate an approach that generates actionable knowledge to alleviate socio-technical gaps.
Our approach takes into account the past history of a project analyzing socio-technical gaps with respect to their individual relation to build failure.
These recommendations aim at initiating coordination between two developers that form a gap in the current build, thus increasing the chance of a build to succeed.
This approach can support projects that use electronic repositories such as automated build engines, version control and task management.

For our approach we define a socio-technical gap as the relationship between two developers that share a technical dependency (implying coordination need) without any social interaction (implying unmet coordination need).
A technical dependency can be inferred by two developers changing the same file, or a developer changing method another developer's code is calling.
%In this paper, two developers that share a technical dependency are referred to as a \emph{technical pair} (or socio-technical gap).
In contrast two developers that share a technical dependency and that also coordinate their work are referred to as being in a \emph{socio-technical pair}.
For example, two developers that discuss their work through email are said to coordinate, and if they additionally share a technical dependency then they form a socio-technical pair.
%
When analyzing the developer pairs in a project's set of builds, in our approach we recognize that each \emph{technical} pair can have a \emph{corresponding socio-technical} if the same two developers have a technical dependency matched by actual coordination in a different build. 

%two developers that are in a technical pair in some builds can also have a \emph{corresponding socio-technical} relationship in a different build (i.e. where their technical dependency is met by actual coordination). 

%We further define a technical and a socio-technical pair to be \emph{corresponding} if both pairs consist of the same pair of developers.

Our approach analyzes the technical pairs in relation to build failure in the following steps:

\begin{enumerate}
\item Identify the set of all technical pairs $T$ across all builds in the project.

\item From set $T$, select the set of \emph{harmful} pairs $H$ by identifying those technical pairs that are statistically related to build failure. To determine the statistical relationship we employ a Fisher exact value test that comparing the frequency of each technical pair's occurrence in failed vs successful builds. The p-values of the Fisher Exact Value test should be adjusted to account for multiple hypothesis testing.

%by counting for each technical pair in how many failed and successful builds it occurs.

\item To form our set of recommendations $R$  we remove from $H$ those pairs where the corresponding socio-technical pair is statistically related to failure ($H_f$) (which would indicate that even matching the technical dependency with actual coordination would not prevent the build from failing). 

Therefore $R = H - H_f$.


%From the set of harmful pairs  (S) we then remove those pairs that have a \emph{corresponding} socio-technical pair that is statistically related to build failure (using a Fisher exact value test as above) forming the set of recommendations.

\item Having identified $R$, we further refine our recommendation by identifying two sets of technical pairs: $R_1$ contains the pairs that have a corresponding socio-technical pair that relates statistically to success and $R_2$ contains the pairs that either (1) have a corresponding socio-technical pair that does not relate to success or failure, or (2) do not have a corresponding socio-technical pair.

Therefore, $R = R_1 \cup R_2$.

In our recommendation set $R_1$ has highest priority because it contains developer pairs that contributed to a successful build when matched by actual coordination.  

%We split the set of recommendations into two sets of recommendations.
%The first set contains those which have a corresponding socio-technical pair that is statistically related to build success (using a Fisher exact value test as mentioned earlier) whereas the second set contains the remaining technical pairs.

\item Finally, for each set in our recommendation $R$ ($R_1$ and $R_2$) rank the developer pairs using the coefficient $p_x$, which represents the normalized likelihood of a build
to fail in the presence of the specific pair:
$$
p_x\text{=}\frac{ \text{pair}_{failed} / \text{total}_{failed} }
                     { \text{pair}_{failed} / \text{total}_{failed} + \text{pair}_{success} / \text{total}_{successs}}
$$
where: pair$_{failed}$ is the number of failed builds in which the pair occurred; total$_{failed}$ is the number of failed builds; pair$_{success}$ is the number of successful builds in which the pair occurred, and total$_{success}$ is the number of successful builds.
A value of $p_x$ closer to one means that the developer pair is strongly related to build
failure. 
\end{enumerate}

In summary, our approach analyzes the technical dependencies, actual coordination and build quality in all existing builds in the project, and recommends a ranked list of developer pairs that, if present in the current build, will increase the current build's chance of failure. This list is prioritized by the probability of this chance. This recommendation essentially represents the pairs of developers that should communicate in order to increase the chance of a build to succeed. In a busy manager's workday, the ranking of each developer pair is useful in prioritizing which socio-technical gaps should be closed first. 

%produces two sets of recommendations ($S_1$ and $S_2$) both containing technical pairs that are related to build failure.
%$S_1$ consists of technical pairs that have an equivalent socio-technical pair that is related to build success.
%$S_2$ consists of technical pairs that do not have an equivalent socio-technical pair that is related to either build outcome.
%Both sets of technical pairs represent recommendations of pairs of developers that should communicate in order to increase the chance of a build to succeed. 
%The technical pairs contained in $S_1$ have a potentially larger influence on increase build success due to converting a build failure related pair into a build success related pair.




\input{eval}



\section{Conclusion and Next Steps}
Motivated by findings in the literature suggesting that socio-technical gaps increase build failure~\cite{kwan:tse:2011},
we hypothesized that these findings can be made actionable.
In this paper we proposed an approach that is capable of generating actionable knowledge that helps developers to prevent builds from failing.

Our results from the preliminary analysis with the RTC data indicate that 
the influence of developer pairs in a socio-technical gap on the build
failure was very high.
This knowledge can be used to improve the chances of the upcoming build to succeed.
Managers that oversee the work of several developers can initiate coordination between developers in technical pairs identified by our approach.
Similarly, a developer that is part of a technical pair in our recommendation can coordinate her work with her colleague to remove the risk posed by this open gap.

% We also found that historical project information about technical pairs and software
%builds can be used in a model that predicts the quality of upcoming builds.
%This means that if any one of the twenty most frequent failure related pairs was present in a social network of a build, the build had at least an 74\% chance to fail.

%PUT MORE

As a next step in our evaluation we have planned three follow up studies:
% MORE HERE be more explicit
(1) We want to examine the extent to which the technical pairs identified by our approach cause builds to fail. 
In our current approach we presented statistical evidence that the technical pairs relate to build failure, as a next step we will examine each technical pair and the failed builds it occurred in to determine if the technical pair actually cause the build to fail.
For example, we plan to identify the changes that broke a build and relate it back to the developers that made the changes and those that were affected by these changes.
%
%HOW CAN IT BE USED

(2) Develop a tool that implements our approach and provides real-time recommendations to the developer, and evaluated it with software practitioners. By comparing teams that do use the recommender system with those teams that do not, we plan on determining the effect the recommendations generated by our approach have on lowering the number of failed builds.

(3) Refine the definition of coordination and its needs, by both leveraging alternative conceptualizations of technical and social dependencies.
For example, we currently consider two developers having a technical dependency if they modified the same file, but plan to investigate alternative conceptualizations such as inferring technical dependencies from who uses code that has been modified.
\bibliographystyle{IEEEtran}
\bibliography{bib}

\end{document}


